From Reversible Programs to \\ Univalent Universes and Back

### 2021-07-07 Pi+/paper/equivalence.tex
Change and fix references

We present the types and 1-combinators of $\PiPlusLang$ and $\PiHatLang$ in~\cref*{fig:piplus,fig:pihat} respectively,
eliding the 2-combinators for brevity. \note{We enforce that there is a unique 2-combinator between compatible
  1-combinators, by introducing a level 3, which is trivial in the same way as in the case of $\PiLang$, as described
  before.} They can be found in the appendix and in the accompanying Agda code.

@@

First, we show how to translate $\PiLang$ programs to $\PiPlusLang$, which is the additive fragment of $\PiLang$. The
syntax is given in~\cref{fig:piplus}.

@@

Next, we show how to translate $\PiPlusLang$ programs to $\PiHatLang$ and back. $\PiHatLang$ is a simplified variant of
$\PiPlusLang$, with (unary) natural numbers for 0-cells, 1-combinators generated by adjacent transpositions, and an
appropriate set of 2-combinators. We give the syntax, again omitting 2-combinators, in~\cref{fig:pihat}.

@@

Types in $\PiHatLang$ are interpreted as 0-cells in $\UFin$, that is, a natural number $n$ is mapped to $\Fin[n]$. The
1-combinators in $\PiHatLang$ are mapped to 1-paths in $\UFin$, that is, 1-loops $\Aut[\Fin[n]]$, in each connected
component. In $\PiHatLang$, the 1-combinators are generated by adjacent transpositions, so these can be easily mapped to
words in $\Sn$ and then to automorphisms using~\cref{prop:sn-lehmer-fin-equiv}. Finally, 2-combinators are mapped to
2-paths between loops in $\UFin$.

@@

The semantics that we presented here takes a different route to constructing the permutation from a $\PiLang$
combinator, compared to the direct interpretation given using the big-step interpreter in~\cref{subsec:denotational}. We
verify that the two semantics agree, establishing that the semantics is adequate and fully abstract.

>>>>

We present the types and 1-combinators of $\PiPlusLang$ and $\PiHatLang$ in~\Cref*{fig:piplus,fig:pihat} respectively,
eliding the 2-combinators for brevity. \note{We enforce that there is a unique 2-combinator between compatible
  1-combinators, by introducing a level 3, which is trivial in the same way as in the case of $\PiLang$, as described
  before.} They can be found in the appendix and in the accompanying Agda code.

@@

First, we show how to translate $\PiLang$ programs to $\PiPlusLang$, which is the additive fragment of $\PiLang$. The
syntax is given in~\Cref{fig:piplus}.

@@

Next, we show how to translate $\PiPlusLang$ programs to $\PiHatLang$ and back. $\PiHatLang$ is a simplified variant of
$\PiPlusLang$, with (unary) natural numbers for 0-cells, 1-combinators generated by adjacent transpositions, and an
appropriate set of 2-combinators. We give the syntax, again omitting 2-combinators, in~\Cref{fig:pihat}.

@@

Types in $\PiHatLang$ are interpreted as 0-cells in $\UFin$, that is, a natural number $n$ is mapped to $\Fin[n]$. The
1-combinators in $\PiHatLang$ are mapped to 1-paths in $\UFin$, that is, 1-loops $\Aut[\Fin[n]]$, in each connected
component. In $\PiHatLang$, the 1-combinators are generated by adjacent transpositions, so these can be easily mapped to
words in $\Sn$ and then to automorphisms using~\Cref{prop:sn-lehmer-fin-equiv}. Finally, 2-combinators are mapped to
2-paths between loops in $\UFin$.

@@

The semantics that we presented here takes a different route to constructing the permutation from a $\PiLang$
combinator, compared to the direct interpretation given using the big-step interpreter in~\Cref{subsec:denotational}. We
verify that the two semantics agree, establishing that the semantics is adequate and fully abstract.

### 2021-07-07 Pi+/paper/finite.tex
Change and fix references

% Recap of the previous section
In~\cref{sec:ufin}, we established that paths in $\UFin$ are equivalent to families of loops on $\Fin[n]$ for every
$n:\Nat$, that is, automorphisms of finite sets of size $n$, with the loopspace encoding the automorphism group. This is
also known to be the finite symmetric group $\Sn[n]$. In the following sections, we will describe this group
syntactically.

@@

% High-level explanation of technical devices
First, we will define the required notions of free groups and group presentations, and state some of their most
important properties. Then, we introduce our chosen \emph{Coxeter presentation} for $\Sn$. To solve the word problem for
$\Sn$, we will use a rewriting system, with a suitable, well behaved collection of reduction rules corresponding to the
Coxeter presentation. Finally, we describe the normal forms in this rewriting system, using Lehmer codes, and prove the
correspondence between them and the type $\Aut[\Fin[n]]$ of automorphisms on a finite set. The generators and relations
we use here will be used to quote back to 1 and 2-combinators in $\PiLang$ (see~\cref{sec:equivalence}).

@@

We use a presentation based on adjacent transpositions for generators, called a~\emph{Coxeter presentation}, which we
also use to solve the word problem for $\Sn$. Connecting it back to the language $\PiLang$, we will show that it is
possible to encode all $\PiLang$ combinators using adjacent transpositions (in~\cref{sec:equivalence}). The rest of this
section develops these intuitions formally. Group presentations are built by adding equations to a \emph{free group},
which we define first.

@@

  From universal algebra, a group is simply a set with a 0-ary constant $e$ (the neutral element), a binary operation
  $\blank\mult\blank$ for group multiplication, and a unary inverse operation $\inv{\blank}$. The neutral element has to
  satisfy unit and inverse laws, and the multiplication has to be associative (see~\cref{def:group}).

@@

\begin{toappendix}
  \begin{definition}
    \label{def:free-group}
    Given a set $A$, the free group $F(A)$ on it is given by a higher inductive type with the following point and path
    constructors. Notice the similarity with the definition of a group structure (\cref{def:group}), but note that each
    operation here is a generator for $F(A)$.
    \begin{itemize}
      \item An inclusion function $\eta_{A} : A \to F(A)$
      \item A multiplication function $m : F(A) \times F(A) \to F(A)$
      \item An element $e : F(A)$
      \item An inverse function $i : F(A) \to F(A)$
    \end{itemize}
    \smallskip
    \begin{itemize}
      \item For every $x, y, z : F(A)$, a path $\term{assoc} : m(x, m(y, z)) \id m(m(x, y), z)$
      \item For every $x : F(A)$, paths $\term{unitr} : m(x, e) \id x$ and $\term{unitl} : m(e, x) \id x$
      \item For every $x : F(A)$, paths $\term{invr} : m(x, i(x)) \id e$ and $\term{invl} : m(i(x), x) \id e$
      \item A 0-truncation, for every $x, y : F(A)$ and $p, q : x \id y$, a 2-path $\term{trunc} : p \id q$
    \end{itemize}
  \end{definition}
\end{toappendix}

@@

Following the universal-algebraic definition, in HoTT, we could use a naive higher inductive type to define the free
group, which enforces the group axioms by adding path constructors (see~\cref{def:free-group}).
Using the induction principle, we can easily verify the universal property. However, since this definition of $F(A)$ has
lots of path constructors corresponding to each group axiom, characterising its path space is difficult.

@@

\begin{proposition}
  $F(A) \defeq \quot{\List[A + A]}{\sim^{\ast}}$ has a group structure, with the empty list $\nil$ for the neutral
  element, multiplication given by list append $\append$, and inverse given by flipping $\inl$ and $\inr$, followed by
  reversing the list. Further, $F(A)$ with $\eta_A(a) \defeq \inl(a) \cons \nil$ satisfies the universal property of
  free groups, as stated in~\cref{prop:free-groups}.
\end{proposition}

@@

The universal property of the above definition is similar to~\cref{prop:free-groups} except the relation has to be
preserved by the function mapping out of the generating set.

@@

To present the group $\Sn$, the primitive operations we are going to use will be adjacent swaps. When dealing with
permutations on an $n + 1$-element set, there are $n$ adjacent transpositions, transposition number $k$ swapping
elements $(\el{k}, \el{k+1})$. Thus, the generating set would be $\Fin[n]$. There are three relations that we're going
to specify for this presentation -- these are the equations that these generators should satisfy. We visualise them as
braid diagrams in~\cref{fig:coxeter-braid}.

@@

This construction is called a Coxeter presentation of $\Sn$. Writing it formally, we encode the rules discussed above
using a relation $\cox$ on $\List[\Fin[n]]$ (\cref{def:cox}), and then take its congruence closure $\cox*$
(\cref{def:coxstar}).

@@

For this to work, we first need the system to have the termination property, meaning that there are no infinite
reductions. We observe that after throwing out reflexivity and symmetry, the right hand sides of the relations $\cox*$
are strictly smaller than the left hand sides, in terms of the lexicographical ordering on words in $\Fin[n]$ (which is
well-founded). Thus, by directing the relation from left to right, we would get the termination property out of the box.
Second, we need the normal forms to be unique, so that we can get a normalisation function. This will be true if the
rewriting system is confluent -- meaning that all critical pairs, that is, terms with overlapping possible reduction
rules, have to converge. For example, in our system, the pairs in~\cref{fig:critical-pairs-converging} converge.
Unfortunately, this is not true for all critical pairs -- an  example is in~\cref{fig:critical-pairs-non-converging},
where both endpoints are normal with respect to the $\cox*$ relation.

@@

The result of this function is the sequence \([k + n - 1, k + n - 2, k + n - 3, \ldots, n]\). Since we think of the
number $m$ as the transposition swapping elements $(\el{m}, \el{m+1})$, the role of this helper function
is to produce a sequence of transpositions -- a permutation -- which moves element $\el{k + n}$ by $k$ places left,
shifting all the elements in between one place right (see~\cref{fig:downf}).

@@

Constructors $\longcancel$ and $\longswap$ correspond directly to the appropriate constructors of $\cox$ and can be
visualised in the same way as before. The remaining constructor uses the helper function to exchange the order of a long
sequence of transpositions and a single transposition afterwards. For example, for $n = 0$ and $k = 3$, it allows for
the reduction $[4, 3, 2, 1, 0, 4] \longcox [3, 4, 3, 2, 1, 0]$ (see~\cref{fig:longcox}).

@@

We define the relation $\longcox*$ to be the closure of $\longcox$, under reflexivity and (right-step extended)
transitivity (see~\cref{def:longcox-star}).

@@

By~\cref{prop:nf}~\ref{prop:nf-uniqueness}, we get a unique choice function $\normf : {\List[\Fin[n]]} \to
  {\List[\Fin[n]]}$ that produces a normal form for terms of $\List[\Fin[n]]$. We state and prove a couple of important
properties enjoyed by $\normf$.

@@

Notice however, that a group presentation, as defined in~\cref{def:presentation}, requires the relation to be on the set
of words $A + A$, where the right copy corresponds to the set of formal inverses of the generators. The constructor
$\cancel$ specifies that the inverse of each element is again the same element, using which we can show that our
definition of $\Sn$ is equivalent to the definition of a presented group, by lifting the $\cox*$ relation.

@@

\begin{proof}
  For $n = \zero$, note that $\Lehmer[\zero]$ is contractible, and so is $\Aut[\Fin[\suc[\zero]]]$. For $n = \suc[m]$,
  we compute a chain of equivalences.
  \begin{gather*}
    \arraycolsep=0.5em\def\arraystretch{1.5}
    \begin{array}{rl}
           & \Lehmer[\zero]          \\
      \eqv & \unit                   \\
      \eqv & \Aut[\Fin[\suc[\zero]]] 
    \end{array}
    \qquad\qquad
    \begin{array}{rlr}
           & \Lehmer[\suc[m]]                                                                                   &                                                                 \\
      \eqv & \Fin[\suc[\suc[m]]] \times \Lehmer[m]                                                              & \text{by definition}                                            \\
      \eqv & \Fin[\suc[\suc[m]]] \times \Aut[\Fin[\suc[m]]]                                                     & \text{induction hypothesis}                                     \\
      \eqv & \dsum{k : \Fin[\suc[\suc[m]]]}{\Fin[\suc[m]] \eqv \Fin[\suc[m]]}                                   & \text{$\Sigma$ over a constant family}                          \\
      \eqv & \dsum{k : \Fin[\suc[\suc[m]]]}{\FinExcept{\suc[\suc[m]]}[m] \eqv \Fin[\suc[m]]}                    & \text{by~\cref{prop:fin-finexcept}~\cref{prop:fin-finexcept-2}} \\
      \eqv & \dsum{k : \Fin[\suc[\suc[m]]]}{\FinExcept{\suc[\suc[m]]}[m] \eqv \FinExcept{\suc[\suc[m]]}[m - k]} & \text{by~\cref{prop:fin-finexcept}~\cref{prop:fin-finexcept-2}} \\
      \eqv & \Aut[\Fin[\suc[\suc[m]]]]                                                                          & \text{by~\cref{prop:fin-finexcept}~\cref{prop:fin-finexcept-3}}
    \end{array}
  \end{gather*}
\end{proof}

By composing~\cref{prop:lehmer-aut-equiv} and~\cref{prop:sn-im-lehmer-equiv}, we obtain the final equivalence.

\begin{corollary}
  For all $n : \Nat$,
  \(
  \Sn \eqv \Lehmer[n] \eqv \Aut[\Fin[\suc[n]]]
  \).
\end{corollary}

>>>>

% Recap of the previous section
In~\Cref{sec:ufin}, we established that paths in $\UFin$ are equivalent to families of loops on $\Fin[n]$ for every
$n:\Nat$, that is, automorphisms of finite sets of size $n$, with the loopspace encoding the automorphism group. This is
also known to be the finite symmetric group $\Sn[n]$. In the following sections, we will describe this group
syntactically.

@@

% High-level explanation of technical devices
First, we will define the required notions of free groups and group presentations, and state some of their most
important properties. Then, we introduce our chosen \emph{Coxeter presentation} for $\Sn$. To solve the word problem for
$\Sn$, we will use a rewriting system, with a suitable, well behaved collection of reduction rules corresponding to the
Coxeter presentation. Finally, we describe the normal forms in this rewriting system, using Lehmer codes, and prove the
correspondence between them and the type $\Aut[\Fin[n]]$ of automorphisms on a finite set. The generators and relations
we use here will be used to quote back to 1 and 2-combinators in $\PiLang$ (see~\Cref{sec:equivalence}).

@@

We use a presentation based on adjacent transpositions for generators, called a~\emph{Coxeter presentation}, which we
also use to solve the word problem for $\Sn$. Connecting it back to the language $\PiLang$, we will show that it is
possible to encode all $\PiLang$ combinators using adjacent transpositions (in~\Cref{sec:equivalence}). The rest of this
section develops these intuitions formally. Group presentations are built by adding equations to a \emph{free group},
which we define first.

@@

  From universal algebra, a group is simply a set with a 0-ary constant $e$ (the neutral element), a binary operation
  $\blank\mult\blank$ for group multiplication, and a unary inverse operation $\inv{\blank}$. The neutral element has to
  satisfy unit and inverse laws, and the multiplication has to be associative (see~\Cref{def:group}).

@@

\begin{toappendix}
  \begin{definition}
    \label{def:free-group}
    Given a set $A$, the free group $F(A)$ on it is given by a higher inductive type with the following point and path
    constructors. Notice the similarity with the definition of a group structure (\Cref{def:group}), but note that each
    operation here is a generator for $F(A)$.
    \begin{itemize}
      \item An inclusion function $\eta_{A} : A \to F(A)$
      \item A multiplication function $m : F(A) \times F(A) \to F(A)$
      \item An element $e : F(A)$
      \item An inverse function $i : F(A) \to F(A)$
    \end{itemize}
    \smallskip
    \begin{itemize}
      \item For every $x, y, z : F(A)$, a path $\term{assoc} : m(x, m(y, z)) \id m(m(x, y), z)$
      \item For every $x : F(A)$, paths $\term{unitr} : m(x, e) \id x$ and $\term{unitl} : m(e, x) \id x$
      \item For every $x : F(A)$, paths $\term{invr} : m(x, i(x)) \id e$ and $\term{invl} : m(i(x), x) \id e$
      \item A 0-truncation, for every $x, y : F(A)$ and $p, q : x \id y$, a 2-path $\term{trunc} : p \id q$
    \end{itemize}
  \end{definition}
\end{toappendix}

@@

Following the universal-algebraic definition, in HoTT, we could use a naive higher inductive type to define the free
group, which enforces the group axioms by adding path constructors (see~\Cref{def:free-group}).
Using the induction principle, we can easily verify the universal property. However, since this definition of $F(A)$ has
lots of path constructors corresponding to each group axiom, characterising its path space is difficult.

@@

\begin{proposition}
  $F(A) \defeq \quot{\List[A + A]}{\sim^{\ast}}$ has a group structure, with the empty list $\nil$ for the neutral
  element, multiplication given by list append $\append$, and inverse given by flipping $\inl$ and $\inr$, followed by
  reversing the list. Further, $F(A)$ with $\eta_A(a) \defeq \inl(a) \cons \nil$ satisfies the universal property of
  free groups, as stated in~\Cref{prop:free-groups}.
\end{proposition}

@@

The universal property of the above definition is similar to~\Cref{prop:free-groups} except the relation has to be
preserved by the function mapping out of the generating set.

@@

To present the group $\Sn$, the primitive operations we are going to use will be adjacent swaps. When dealing with
permutations on an $n + 1$-element set, there are $n$ adjacent transpositions, transposition number $k$ swapping
elements $(\el{k}, \el{k+1})$. Thus, the generating set would be $\Fin[n]$. There are three relations that we're going
to specify for this presentation -- these are the equations that these generators should satisfy. We visualise them as
braid diagrams in~\Cref{fig:coxeter-braid}.

@@

This construction is called a Coxeter presentation of $\Sn$. Writing it formally, we encode the rules discussed above
using a relation $\cox$ on $\List[\Fin[n]]$ (\Cref{def:cox}), and then take its congruence closure $\cox*$
(\Cref{def:coxstar}).

@@

For this to work, we first need the system to have the termination property, meaning that there are no infinite
reductions. We observe that after throwing out reflexivity and symmetry, the right hand sides of the relations $\cox*$
are strictly smaller than the left hand sides, in terms of the lexicographical ordering on words in $\Fin[n]$ (which is
well-founded). Thus, by directing the relation from left to right, we would get the termination property out of the box.
Second, we need the normal forms to be unique, so that we can get a normalisation function. This will be true if the
rewriting system is confluent -- meaning that all critical pairs, that is, terms with overlapping possible reduction
rules, have to converge. For example, in our system, the pairs in~\Cref{fig:critical-pairs-converging} converge.
Unfortunately, this is not true for all critical pairs -- an  example is in~\Cref{fig:critical-pairs-non-converging},
where both endpoints are normal with respect to the $\cox*$ relation.

@@

The result of this function is the sequence \([k + n - 1, k + n - 2, k + n - 3, \ldots, n]\). Since we think of the
number $m$ as the transposition swapping elements $(\el{m}, \el{m+1})$, the role of this helper function
is to produce a sequence of transpositions -- a permutation -- which moves element $\el{k + n}$ by $k$ places left,
shifting all the elements in between one place right (see~\Cref{fig:downf}).

@@

Constructors $\longcancel$ and $\longswap$ correspond directly to the appropriate constructors of $\cox$ and can be
visualised in the same way as before. The remaining constructor uses the helper function to exchange the order of a long
sequence of transpositions and a single transposition afterwards. For example, for $n = 0$ and $k = 3$, it allows for
the reduction $[4, 3, 2, 1, 0, 4] \longcox [3, 4, 3, 2, 1, 0]$ (see~\Cref{fig:longcox}).

@@

We define the relation $\longcox*$ to be the closure of $\longcox$, under reflexivity and (right-step extended)
transitivity (see~\Cref{def:longcox-star}).

@@

By~\Cref{prop:nf}~\ref{prop:nf-uniqueness}, we get a unique choice function $\normf : {\List[\Fin[n]]} \to
  {\List[\Fin[n]]}$ that produces a normal form for terms of $\List[\Fin[n]]$. We state and prove a couple of important
properties enjoyed by $\normf$.

@@

Notice however, that a group presentation, as defined in~\Cref{def:presentation}, requires the relation to be on the set
of words $A + A$, where the right copy corresponds to the set of formal inverses of the generators. The constructor
$\cancel$ specifies that the inverse of each element is again the same element, using which we can show that our
definition of $\Sn$ is equivalent to the definition of a presented group, by lifting the $\cox*$ relation.

@@

\begin{proof}
  For $n = \zero$, note that $\Lehmer[\zero]$ is contractible, and so is $\Aut[\Fin[\suc[\zero]]]$. For $n = \suc[m]$,
  we compute a chain of equivalences.
  \begin{gather*}
    \arraycolsep=0.5em\def\arraystretch{1.5}
    \begin{array}{rl}
           & \Lehmer[\zero]          \\
      \eqv & \unit                   \\
      \eqv & \Aut[\Fin[\suc[\zero]]]
    \end{array}
    \qquad\qquad
    \begin{array}{rlr}
           & \Lehmer[\suc[m]]                                                                                   &                                                                 \\
      \eqv & \Fin[\suc[\suc[m]]] \times \Lehmer[m]                                                              & \text{by definition}                                            \\
      \eqv & \Fin[\suc[\suc[m]]] \times \Aut[\Fin[\suc[m]]]                                                     & \text{induction hypothesis}                                     \\
      \eqv & \dsum{k : \Fin[\suc[\suc[m]]]}{\Fin[\suc[m]] \eqv \Fin[\suc[m]]}                                   & \text{$\Sigma$ over a constant family}                          \\
      \eqv & \dsum{k : \Fin[\suc[\suc[m]]]}{\FinExcept{\suc[\suc[m]]}[m] \eqv \Fin[\suc[m]]}                    & \text{by~\Cref{prop:fin-finexcept}~\Cref{prop:fin-finexcept-2}} \\
      \eqv & \dsum{k : \Fin[\suc[\suc[m]]]}{\FinExcept{\suc[\suc[m]]}[m] \eqv \FinExcept{\suc[\suc[m]]}[m - k]} & \text{by~\Cref{prop:fin-finexcept}~\Cref{prop:fin-finexcept-2}} \\
      \eqv & \Aut[\Fin[\suc[\suc[m]]]]                                                                          & \text{by~\Cref{prop:fin-finexcept}~\Cref{prop:fin-finexcept-3}}
    \end{array}
  \end{gather*}
\end{proof}

By composing~\Cref{prop:lehmer-aut-equiv} and~\Cref{prop:sn-im-lehmer-equiv}, we obtain the final equivalence.

\begin{corollary}
  \label{prop:sn-lehmer-fin-equiv}
  For all $n : \Nat$,
  \(
  \Sn \eqv \Lehmer[n] \eqv \Aut[\Fin[\suc[n]]]
  \).
\end{corollary}

### 2021-07-07 Pi+/paper/ufin.tex
Change and fix references

% Fourth - expain the tool we use for fixing it
Every type in HoTT is a weak $\infty$-groupoid whose points are the terms of the type, and the (iterated) identity type
gives the (higher) morphisms. The groupoid we're interested in has types as points, type equivalences for 1-cells, and
higher homotopies for higher cells. This is the groupoid structure for the universe type $\UU$, since the identity type
on types can be characterised as type equivalences (by \emph{univalence}). But, we only want to carve out a
\emph{subuniverse} of \emph{finite types}, still satisfying univalence, to get the groupoid structure. In this section,
we formally define \emph{univalent subuniverses}, and proceed to construct the particular instance for finite types,
$\UFin$ (\cref{def:ufin}).

@@

  \note{We use the rig structure of $\UFin$ in~\cref{subsec:rig} to interpret $\PiLang$.}

@@

  Given two terms $x:A$ and $y:A$, we write $x \id_{A} y$, or simply $x \id y$, for the identity type, which is the type
  of equalities or identifications between them. The identity type is generated by reflexivity $\refl_{x} : x \id_{A} x$,
  and the eliminator for the identity type is given by path induction or the $J$-rule (\cref{def:path-induction}). This
  construction can be iterated, giving the identity type between two terms of an identity type, repeating ad infinitum.
  Using the iterated identity type for morphisms, each type is equipped with the structure of a weak $\infty$-groupoid,
  where each morphism satisfies groupoid laws only upto a higher one. Given an arbitrary type (or groupoid) $A$, we list
  some laws that are provable using path induction.

@@

  The type $\dsum{x:A}{P(x)}$ is the collection of all the fibers and is called the total space of $P$. The first
  projection ${\pi_1 : \dsum{x:A}{P(x)} \to A}$ from the total space to the base space $A$ has the structure of a
  fibration, that is, there is a lifting operation (\cref{fig:lift}) which lifts paths in the base space to paths in the
  total space. Given a path $p : x \id_{A} y$ in the base space, and $u : P(x)$ a point in the fiber over $x$, we have

@@

  Given a type $A$, the propositional truncation $\Trunc[-1]{A}$, squashes the elements of $A$ turning it into a
  proposition. It is given by a HIT with a point constructor $\trunc{\blank} : A \to \Trunc{A}$, and a path constructor
  $\term{trunc}(x,y) : x \id_{\Trunc{A}} y$, which equates every pair of points in the truncation
  (see~\cref{def:prop-trunc}).

@@

  Another HIT that we use is the set-quotient $\quot{A}{R}$ which takes an $\hSet$ $A$ and a relation $R : A \to A \to
    \UU$. It has an inclusion of points $q : A \to \quot{A}{R}$, and adds paths between related pairs of elements
  $\quotrel : R(x,y) \to q(x) \id_{\quot{A}{R}} q(y)$ (see~\cref{def:set-quot}).  We recall that the quotient is
  \emph{effective} if $R$ is a prop-valued equivalence relation, that is, $R(x,y)$ holds iff $(q(x) \id_{\quot{A}{R}}
    q(y))$.

@@

The types we're interested in are the finite types. In constructive mathematics, the notion of finiteness is
subtle~\cite{spiwackConstructivelyFinite2010}. We use the notion of Bishop-finiteness -- a type is finite if it is
merely equivalent to a finite set (\cref{def:finite-set}), and we define the predicate $\isFin$ (\cref{def:isfin}) to do
so.

@@

\begin{proposition}
  For any $n, m : \Nat$,
  \begin{gather*}
    \begin{aligned}
      \Fin[0]                & \eqv \bot        \\
      \Fin[n] \sqcup \Fin[m] & \eqv \Fin[n + m] \\
    \end{aligned}
    \qquad
    \begin{aligned}
      \Fin[1]                & \eqv \top        \\
      \Fin[n] \times \Fin[m] & \eqv \Fin[n * m] \\
    \end{aligned}
  \end{gather*}
  and for any types $X, Y, Z$,
  \begin{gather*}
    \begin{aligned}
      \bot \sqcup X         & \eqv X                     \\
      X \sqcup \bot         & \eqv X                     \\
      (X \sqcup Y) \sqcup Z & \eqv X \sqcup (Y \sqcup Z) \\
      X \sqcup Y            & \eqv Y \sqcup Y            \\
    \end{aligned}
    \qquad
    \begin{aligned}
      \top \times X         & \eqv X                                \\
      X \times \top         & \eqv X                                \\
      (X \times Y) \times Z & \eqv X \times (Y \times Z)            \\
      X \times Y            & \eqv Y \times X                       \\
      X \times \bot         & \eqv \bot                             \\
      X \times (Y \sqcup Z) & \eqv (X \times Y) \sqcup (X \times Z)
    \end{aligned}
  \end{gather*}
\end{proposition}

>>>>

% Fourth - expain the tool we use for fixing it
Every type in HoTT is a weak $\infty$-groupoid whose points are the terms of the type, and the (iterated) identity type
gives the (higher) morphisms. The groupoid we're interested in has types as points, type equivalences for 1-cells, and
higher homotopies for higher cells. This is the groupoid structure for the universe type $\UU$, since the identity type
on types can be characterised as type equivalences (by \emph{univalence}). But, we only want to carve out a
\emph{subuniverse} of \emph{finite types}, still satisfying univalence, to get the groupoid structure. In this section,
we formally define \emph{univalent subuniverses}, and proceed to construct the particular instance for finite types,
$\UFin$ (\Cref{def:ufin}).

@@

  \note{We use the rig structure of $\UFin$ in~\Cref{subsec:rig} to interpret $\PiLang$.}

@@

  Given two terms $x:A$ and $y:A$, we write $x \id_{A} y$, or simply $x \id y$, for the identity type, which is the type
  of equalities or identifications between them. The identity type is generated by reflexivity $\refl_{x} : x \id_{A} x$,
  and the eliminator for the identity type is given by path induction or the $J$-rule (\Cref{def:path-induction}). This
  construction can be iterated, giving the identity type between two terms of an identity type, repeating ad infinitum.
  Using the iterated identity type for morphisms, each type is equipped with the structure of a weak $\infty$-groupoid,
  where each morphism satisfies groupoid laws only upto a higher one. Given an arbitrary type (or groupoid) $A$, we list
  some laws that are provable using path induction.

@@

  The type $\dsum{x:A}{P(x)}$ is the collection of all the fibers and is called the total space of $P$. The first
  projection ${\pi_1 : \dsum{x:A}{P(x)} \to A}$ from the total space to the base space $A$ has the structure of a
  fibration, that is, there is a lifting operation (\Cref{fig:lift}) which lifts paths in the base space to paths in the
  total space. Given a path $p : x \id_{A} y$ in the base space, and $u : P(x)$ a point in the fiber over $x$, we have

@@

  Given a type $A$, the propositional truncation $\Trunc[-1]{A}$, squashes the elements of $A$ turning it into a
  proposition. It is given by a HIT with a point constructor $\trunc{\blank} : A \to \Trunc{A}$, and a path constructor
  $\term{trunc}(x,y) : x \id_{\Trunc{A}} y$, which equates every pair of points in the truncation
  (see~\Cref{def:prop-trunc}).

@@

  Another HIT that we use is the set-quotient $\quot{A}{R}$ which takes an $\hSet$ $A$ and a relation $R : A \to A \to
    \UU$. It has an inclusion of points $q : A \to \quot{A}{R}$, and adds paths between related pairs of elements
  $\quotrel : R(x,y) \to q(x) \id_{\quot{A}{R}} q(y)$ (see~\Cref{def:set-quot}).  We recall that the quotient is
  \emph{effective} if $R$ is a prop-valued equivalence relation, that is, $R(x,y)$ holds iff $(q(x) \id_{\quot{A}{R}}
    q(y))$.

@@

The types we're interested in are the finite types. In constructive mathematics, the notion of finiteness is
subtle~\cite{spiwackConstructivelyFinite2010}. We use the notion of Bishop-finiteness -- a type is finite if it is
merely equivalent to a finite set (\Cref{def:finite-set}), and we define the predicate $\isFin$ (\Cref{def:isfin}) to do
so.

@@

\begin{proposition}
  For any $n, m : \Nat$,
  \begin{gather*}
    \begin{aligned}
      \Fin[0]                & \eqv \bot        \\
      \Fin[n] \sqcup \Fin[m] & \eqv \Fin[n + m]
    \end{aligned}
    \qquad
    \begin{aligned}
      \Fin[1]                & \eqv \top        \\
      \Fin[n] \times \Fin[m] & \eqv \Fin[n * m]
    \end{aligned}
  \end{gather*}
  and for any types $X, Y, Z$,
  \begin{gather*}
    \begin{aligned}
      \bot \sqcup X         & \eqv X                     \\
      X \sqcup \bot         & \eqv X                     \\
      (X \sqcup Y) \sqcup Z & \eqv X \sqcup (Y \sqcup Z) \\
      X \sqcup Y            & \eqv Y \sqcup Y
    \end{aligned}
    \qquad
    \begin{aligned}
      \top \times X         & \eqv X                                \\
      X \times \top         & \eqv X                                \\
      (X \times Y) \times Z & \eqv X \times (Y \times Z)            \\
      X \times Y            & \eqv Y \times X                       \\
      X \times \bot         & \eqv \bot                             \\
      X \times (Y \sqcup Z) & \eqv (X \times Y) \sqcup (X \times Z)
    \end{aligned}
  \end{gather*}
\end{proposition}
